Visual Analysis of Ligand Trajectories in Molecular Dynamics

### 2018-09-27 paper.tex

\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
\usepackage{textcomp}                  % use better special symbols
\usepackage{mathptmx}                  % use matching math font
\usepackage{times}                     % we use Times as the main font
\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
\usepackage{cite}                      % needed to automatically sort the references
\usepackage{tabu}                      % only used for the table example
\usepackage{booktabs}                  % only used for the table example
%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

@@

%% \section{Introduction} %for journal use above \firstsection{..} instead
\begin{itemize}
  \item Motivation
  \item Ligand binding (Method + Trajectory data)
  \item Contribution
\end{itemize}

>>>>

\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
\usepackage{textcomp}                  % use better special symbols
\usepackage{mathptmx}                  % use matching math font
\usepackage{times}                     % we use Times as the main font
\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
\usepackage{cite}                      % needed to automatically sort the references
\usepackage{tabu}                      % only used for the table example
\usepackage{booktabs}                  % only used for the table example
\usepackage{color}
%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

@@

%% \section{Introduction} %for journal use above \firstsection{..} instead

Proteins are essential parts of living organisms. 
Understanding the function of the organisms and their malfunction therefore starts with deep understanding of protein structure and function.
The reactivity of proteins with other molecules is the key property which enables to create new molecules as products of chemical reactions. 
These molecules can serve for different purposes, e.g., can serve as the basis of new drugs. 
In protein engineering, the researchers are focusing on the properties of proteins themselves and the chemical reactions are designed in order to modify these properties. 
An example can be changing the stability of proteins in normal conditions, outside the lab environment~\cite{koudelakova}, or increasing the protein activity towards other molecules~\cite{pavlova}.

Enzymes are forming a specific group of proteins, which accelerate chemical reactions. 
Their have an irreplaceable function in organisms as the enzyme catalysis plays a role in almost all metabolic processes in cells.
The important property of enzymes is that their reaction rate can be increased by lowering their activation energy.
Therefore, studying the energetic profile of the reaction of a small molecule, called ligand, with the enzyme gives the biochemists the information about the feasibility of this reaction to happen.

In other words, when the biochemists are simulating (i.e., modeling) the reaction between the enzyme and ligand, they are using in-silico methods for generating  possible trajectories of the ligand to the enzyme deeply buried site, called the active site.
In this specific site the chemical reaction between the enzyme and ligand can undergo. 
The existing computational methods are generating large ensembles of such trajectories which have to be further explored by the biochemists who are assessing their relevance.
To do this, they have to focus in different properties of the trajectories, including its energetic profile. 
Currently the exploration of possible trajectories is performed by a specific method designed by the biochemists especially for the particular case~\cite{pmid29858243}.
However, a general approach is still missing.
The straightforward solution is designing visualization and visual analysis methods for exploration of trajectories and their properties, enabling the user to interactively filter them according to interesting geometric and physico-chemical aspects.
In this paper, we are presenting a novel tool for visual analysis of large ensembles of possible ligand trajectories, whose design was driven by the actual needs of the protein engineers.
They were also tightly collaborating on the design of the tool, in order to address the needs as much as possible. 
Our tool consists of several linked views enabling the user to explore crucial properties of large ensembles of ligand trajectories, containing thousands of them. 
These can be generated by different simulation algorithms and tools, such as \textcolor{red}{TODO cite caverdock and Vojta's method}.
%~\cite{caverdock,vojta}.

The ultimate goal of our tool is to enable the biochemists to get insight into the relevance of trajectories even when they do not have any apriori knowledge about the trajectories and how to design a method for their processing in a traditionally used manner.

\textcolor{red}{TODO finish that}

%Motivace:
%\begin{itemize}
%  \item Proteiny jsou essential to all living...
%  \item Specificky druh jsou enzymy... metabolismus?
%  \item Dnes chemici bezne zkoumaji funkci enzymu pomoci modelovani, predevsim pomoci molekularni simulace.
%  \item Vysledky in-sillico experimentu jsou objemne a je bezne je analyzovat predem navrzenou metodou s zadnou ci malou znalosti zkoumane simulace [https://loschmidt.chemi.muni.cz/peg/publications/conformational-changes-allow-processing-of-bulky-substrates-by-a-haloalkane-dehalogenase-with-a-small-and-buried-active-site/].
%  \item Cilem naseho vizualizacniho nastroje je umoznit interaktivne "zjistit, co se v datech deje", s malou ci zadnou predstavou o metode jejich zpracovani.
%  Myslim to tak, ze metodu zpracovani si chemik bude moci navrhnout/zpresnit zaroven se samotnou analyzou dat.
%\end{itemize}

Dalsi:
\begin{itemize}
  \item Ligand binding (Method + Trajectory data)
  \item Contribution
\end{itemize}
